Constructing a universe for the setoid model

### 2021-01-16 paper.tex
added Palmgren and small fixes

\author{Thorsten Altenkirch\inst{1}
  \thanks{Supported by USAF grant FA9550-16-1-0029.}
  \orcidID{0000-0002-6582-5025} \and
  Simon Boulier\inst{2}\thanks{Supported by ERC Starting Grant CoqHoTT 637339.} \and
  Ambrus Kaposi\inst{3}\orcidID{0000-0001-9897-8936}
  \thanks{Supported by the National Research, Development and Innovation Fund of Hungary, financed under the Thematic Excellence Programme funding scheme, Project no. ED\_18-1-2019-0030 and the Bolyai Fellowship of the Hungarian Academy of Sciences, Project no.\ BO/00659/19/3.}\and
  Christian Sattler\inst{4}\thanks{Supported by USAF grant FA9550-16-1-0029 and Swedish Research Council grant 2019-03765.} \and
  Filippo Sestini\inst{1}\orcidID{0000-0002-8701-5613}}

\institute{School of Computer Science, University of Nottingham, UK
  \email{\{psztxa,psxfs5\}@nottingham.ac.uk} \and
  Inria, Nantes, France\email{simon.boulier@inria.fr} \and
  E{\"o}tv{\"o}s Lor{\'a}nd University, Budapest, Hungary
  \email{akaposi@inf.elte.hu} \and
  Chalmers University of Technology, Gothenburg, Sweden
  \email{sattler@chalmers.se}
}

@@

The setoid model was first described in \cite{hofmann} in order to add
extensionality principles to Type Theory such as function extensionality and
propositional extensionality. A strict variant of the setoid model was given in
\cite{setoid99} using a definitionally proof-irrelevant universe of
propositions. Recently, support for such a universe was added to the
proof-assistants Agda and Coq \cite{agda-prop}, allowing a full formalization of
Altenkirch's setoid model. Setoid Type Theory (SeTT) is a recently developed
formal system derived from this model construction \cite{mpc19}. Observational
Type Theory (OTT) \cite{alti:ott-conf,alti:ott-conf} is a syntax for the setoid
model differing from SeTT for the use of a different notion of heterogeneous
equality. Moreover, the consistency proof for OTT relies on Extensional Type
Theory, whereas for SeTT it is obtained via a syntactic translation.
%
XTT \cite{xtt} is a cubical variant of OTT where the equality type is defined
using an interval pretype. XTT's universes support universe induction, whereas
it is left open whether the construction presented here supports this principle.
%
Palmgren and Wilander \cite{2014arXiv1408.1364P} construct a setoid
universe using a translation into constructive set theory. In
contrast, we target a basic intensional type theory and hence also
give a computational semantics.

>>>>

\author{Thorsten Altenkirch\inst{1}
  \thanks{Supported by USAF grant FA9550-16-1-0029.}
  \orcidID{0000-0002-6582-5025} \and
  Simon Boulier\inst{2}\thanks{Supported by ERC Starting Grant CoqHoTT 637339.} \and
  Ambrus Kaposi\inst{3}\orcidID{0000-0001-9897-8936}
  \thanks{Supported by the Bolyai Fellowship of the Hungarian Academy
    of Sciences (BO/00659/19/3) and by the ``Application Domain
    Specific Highly Reliable IT Solutions'' project that has been
    implemented with support from the National Research, Development
    and Innovation Fund of Hungary, financed under the Thematic
    Excellence Programme TKP2020-NKA-06 funding scheme.}\and
  Christian Sattler\inst{4}\thanks{Supported by USAF grant FA9550-16-1-0029 and Swedish Research Council grant 2019-03765.} \and
  Filippo Sestini\inst{1}\orcidID{0000-0002-8701-5613}}

\institute{School of Computer Science, University of Nottingham, UK
  \email{\{psztxa,psxfs5\}@nottingham.ac.uk} \and
  Inria, Nantes, France \email{simon.boulier@inria.fr} \and
  E{\"o}tv{\"o}s Lor{\'a}nd University, Budapest, Hungary
  \email{akaposi@inf.elte.hu} \and
  Chalmers University of Technology, Gothenburg, Sweden
  \email{sattler@chalmers.se}
}

@@

The setoid model was first described in \cite{hofmann} in order to add
extensionality principles to Type Theory such as function extensionality and
propositional extensionality. A strict variant of the setoid model was given in
\cite{setoid99} using a definitionally proof-irrelevant universe of
propositions. Recently, support for such a universe was added to the
proof-assistants Agda and Coq \cite{agda-prop}, allowing a full formalization of
Altenkirch's setoid model. Setoid Type Theory (SeTT) is a recently developed
formal system derived from this model construction \cite{mpc19}. Observational
Type Theory (OTT) \cite{alti:ott-conf} is a syntax for the setoid
model differing from SeTT in the use of a different notion of heterogeneous
equality. Moreover, the consistency proof for OTT relies on Extensional Type
Theory, whereas for SeTT it is obtained via a syntactic translation.
%
XTT \cite{xtt} is a cubical variant of OTT where the equality type is defined
using an interval pretype. XTT's universes support universe induction, whereas
it is left open whether the construction presented here supports this principle.
%
Palmgren and Wilander \cite{2014arXiv1408.1364P} construct a setoid
universe using a translation into constructive set theory. Palmgren
\cite{2019arXiv190901414P} constructs an encoding of ETT in ITT
through Aczel's encoding of set theory in type theory
\cite{ACZEL197855}. He uses type theory as a language for his
formalisation but his construction is set-theoretic in nature. Setoids
are utilized to encode sets as arbitrarily branching well-founded
trees quotiented by bisimulation. His notion of family of setoids does
not use strict propositions and it has a weaker form of proof
irrelevance which seems to be not enough to obtain a model of SeTT.
